Mergeable Types

### 2017-02-27 paper/dali.tex
Minor edit
edits
Sec. 2 -> Sec. 2 + Sec. 3

% Formatting commands
% -------------------
\newcommand{\code}[1]{\,{\tt #1}\,}
\newcommand{\spc}[0]{\quad}
\newcommand{\ALT}{~\mid~}
\newcommand{\rel}[1]{{R}_{\mathit{#1}}}
\newcommand{\conj}{~\wedge~}
\newcommand{\disj}{~\vee~}
\newcommand{\rulelabel}[1]{\textrm{\sc {#1}}}
\newcommand{\ilrulelabel}[1]{{\sc #1}}
\newcommand{\RULE}[2]{\frac{\begin{array}{c}#1\end{array}}
                           {\begin{array}{c}#2\end{array}}}
\newcommand{\denot}[1]{\llbracket #1 \rrbracket}
\newcommand{\bind}{>\!\!>\!\!=}
\newcommand{\coloneqq}{::=}
\newcommand{\stepsto}{\longrightarrow}
\newcommand{\return}[1]{\C{return}\;#1}
\newcommand{\run}[2]{\C{run}\;#1\;#2}
\newcommand{\fork}[1]{\C{fork}\;#1}
\newcommand{\push}[1]{\C{push}\;#1}
\newcommand{\pull}{\C{pull}}
\newcommand{\semsucceq}{\succeq_{\circ}}
\newcommand{\sempreceq}{\preceq_{\circ}}
\newcommand{\under}[2]{#1\,\vdash\,#2}
\newcommand{\mbleto}{\rightsquigarrow_m}
\newcommand{\drawsome}{{\sf Canvas}\xspace}
\newcommand{\lang}{\lambda_b}
\newcommand{\length}[1]{\left|#1\right|}

@@

% Listings Code
%---------------
\newcommand{\lstml}{
\lstset{ %
language=ML, % choose the language of the code
basicstyle=\footnotesize\ttfamily,       % the size of the fonts that are used for the code
keywordstyle=\color{Bittersweet},
% numbers=left,                   % where to put the line-numbers
numberstyle=\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
% frame=single,                   % adds a frame around the code
tabsize=2,                      % sets default tabsize to 2 spaces
captionpos=b,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
commentstyle=\itshape\color{MidnightBlue},
%escapeinside={\%*}{*)},         % if you want to add a comment within your code
morekeywords={module, when, @@deriving, not, : , /\\}
}}
\lstnewenvironment{ocaml}
    { % \centering
			\lstml
      \lstset{}%
      \csname lst@setfirstlabel\endcsname}
    { %\centering
      \csname lst@savefirstlabel\endcsname}

@@

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Distributed applications often replicate data to improve
  availability and fault tolerance.  But, the use of replication leads
  to a significantly more complex programming model.  To ensure a
  sensible measure of consistency among different replicas of an
  object, applications are typically structured so that the operations
  applied to a replicated object on one node are subsequently
  communicated to other nodes where they can be re-applied locally.
  Oftentimes, the degree of restructuring and care necessary to
  transform a sequential or shared-memory concurrent application to a
  meaningful (replicated) distributed one can be substantial.

  In this paper, we present a new interpretation of replicated types
  that overcome these concerns.  Our declarative programming model
  called \name\ treats replication as a form of versioning, in which
  consistency is enforced through programmer-supplied \emph{merge}
  functions.  Distributed state is now viewed in terms of a tree of
  immutable object versions, with merge actions used explicitly by the
  application or implicitly by the underlying runtime system to
  periodically reconcile local versions to yield a consistent global
  state.  Thus, our model frees the programmer from having to think of
  low-level operational notions related to distribution and
  replication, and therefore does not entail any restructuring of
  application logic to enable distributed programming.

  We show how any ML datatype can be enriched with mergeable
  semantics, formalize a rich concurrency model based on such types,
  and describe an instantiation of OCaml equipped with these features.
  Our implementation is integrated with a content-addressable
  distributed storage layer that allows any OCaml program to be easily
  transformed into an efficient, highly available, distributed
  application.
\end{abstract}

@@

@@

\input{mergeable-types}

@@

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

>>>>

% Formatting commands
% -------------------
\newcommand{\code}[1]{\,{\tt #1}\,}
\newcommand{\spc}[0]{\quad}
\newcommand{\ALT}{~\mid~}
\newcommand{\rel}[1]{{R}_{\mathit{#1}}}
\newcommand{\conj}{~\wedge~}
\newcommand{\disj}{~\vee~}
\newcommand{\rulelabel}[1]{\textrm{\sc {#1}}}
\newcommand{\ilrulelabel}[1]{{\sc #1}}
\newcommand{\RULE}[2]{\frac{\begin{array}{c}#1\end{array}}
                           {\begin{array}{c}#2\end{array}}}
\newcommand{\denot}[1]{\llbracket #1 \rrbracket}
\newcommand{\bind}{>\!\!>\!\!=}
\ifx\coloneqq\undefined
  \newcommand{\coloneqq}{::=}
\else
  \renewcommand{\coloneqq}{::=}
\fi
\newcommand{\stepsto}{\longrightarrow}
\newcommand{\return}[1]{\C{return}\;#1}
\newcommand{\run}[2]{\C{run}\;#1\;#2}
\newcommand{\fork}[1]{\C{fork}\;#1}
\newcommand{\push}[1]{\C{push}\;#1}
\newcommand{\pull}{\C{pull}}
\newcommand{\semsucceq}{\succeq_{\circ}}
\newcommand{\sempreceq}{\preceq_{\circ}}
\newcommand{\under}[2]{#1\,\vdash\,#2}
\newcommand{\mbleto}{\rightsquigarrow_m}
\newcommand{\drawsome}{{\sf Canvas}\xspace}
\newcommand{\buylist}{{\sf ToBuy}\xspace}
\newcommand{\lang}{\lambda_b}
\newcommand{\length}[1]{\left|#1\right|}
\newcommand{\GK}[1]{\textcolor{red}{GK: #1}}

@@

% Listings Code
%---------------
\newcommand{\lstml}{
\lstset{ %
language=ML, % choose the language of the code
basicstyle=\footnotesize\ttfamily,       % the size of the fonts that are used for the code
keywordstyle=\color{Bittersweet},
% numbers=left,                   % where to put the line-numbers
numberstyle=\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
% frame=single,                   % adds a frame around the code
tabsize=2,                      % sets default tabsize to 2 spaces
captionpos=b,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
commentstyle=\itshape\color{MidnightBlue},
%escapeinside={\%*}{*)},         % if you want to add a comment within your code
morekeywords={module, match, when, @@deriving, not, : , /\\}
}}
\lstnewenvironment{ocaml}
    { % \centering
			\lstml
      \lstset{}%
      \csname lst@setfirstlabel\endcsname}
    { %\centering
      \csname lst@savefirstlabel\endcsname}

@@

\newcommand{\KC}[1]{{\textcolor{red}{{KC: #1}}}}

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Distributed applications often replicate data to improve
  availability and fault tolerance.  But, the use of replication leads
  to a significantly more complex and onerous programming model.  To
  ensure a sensible measure of consistency among different replicas of
  an object, applications are typically structured so that the
  operations applied to a replicated object on one node are
  subsequently communicated to other nodes where they can be
  re-applied locally.  Oftentimes, the degree of restructuring and
  care necessary to transform a sequential or shared-memory concurrent
  application to a meaningful (replicated) distributed one can be
  substantial.

  In this paper, we present a new interpretation of replicated types
  that overcome these concerns.  Our declarative programming model
  called \name\ treats replication as a form of versioning, in which
  consistency is enforced through programmer-supplied \emph{merge}
  functions.  Distributed state is now viewed in terms of a tree of
  immutable object versions, with merge actions used explicitly by the
  application or implicitly by the underlying runtime system to
  periodically reconcile local versions to yield a consistent global
  state.  Our model frees the programmer from having to think of
  low-level operational notions related to distribution and
  replication, and therefore does not entail any restructuring of
  application logic to enable distributed programming.

  We show how any ML datatype can be enriched to support mergeability,
  formalize a concurrency semantics based on such types, and describe
  an instantiation of OCaml equipped with these features.  Our
  implementation is integrated with a content-addressable distributed
  storage layer that allows any OCaml program to be easily transformed
  into an efficient, highly available, distributed application.
\end{abstract}

@@

\input{examples}

@@

\input{system-model}

\input{evaluation}

@@

%% Appendix
\appendix
\input{appendix}
